---
title: "behavior_mod"
author: "Amy Van Scoyoc"
date: "8/14/2022"
output: github_document
---

```{r}
library(here) #reproducible working directory
library(tidyverse) #data wrangling
library(sp) #for spatial points
library(sf) #for spatial transformations with pipes
library(adehabitatHR) #for calculating homerange KUDs
library(raster) #for importing and extracting covariates
library(lme4) #for glmer models
library(performance) #check models
library(jtools) #plot estimates
```

```{r}
#import dataset
load(here("data_objects","coyote.Rdata"))
#import covariates
load(here("data_objects","envt.Rdata"))
```

### Resting top models

### Calculate Rest Home Ranges (Kernel Utilization Distribution)
```{r}
#create object for calculating HRs
coyote_sp <- coyote %>%
   filter(states == 1) %>% #filter to day only
   dplyr::select(ID,x,y) %>% #select data 
   st_as_sf(., coords = c("x","y"), crs = "+init=epsg:4326") %>%  #make spatial object
   st_transform("+init=epsg:26910") %>% #transform back to UTMs
   as("Spatial") #transform back to sp object for KUDs

#create Kernel Utilization Distributions 
kud_all <- adehabitatHR::kernelUD(coyote_sp, h = 'href')
coyote_hr <- adehabitatHR::getverticeshr(kud_all, percent = 95 , unout = "km2")

#HR area (km2)
kernel.area(kud_all, percent = 95) %>% 
  pivot_longer(1:8,names_to = "ID", values_to = "Area (km^2)")
```


### Sample available points in Home Ranges
```{r}
coyote_rest <- coyote %>% 
  filter(states == 1) %>% #filter to rest only
  mutate(Used = 1) %>% #add column to indicate used points
   st_as_sf(., coords = c("x","y"), crs = "+init=epsg:4326") %>%  #make spatial object
   st_transform("+init=epsg:26910")  #transform back to UTMs

coyote_ids <- unique(coyote_rest$ID) # Extract a list of IDs
availables <- list() #empty list to store available pts by ID

# Randomly sample available points from within the home range of each individual
for(i in 1:length(coyote_ids)){
    st_sample(st_as_sf(coyote_hr)[i,], 5*nrow(filter(coyote_rest, ID == coyote_ids[i]))) %>%  #5:1 ratio
    st_sf(geometry = .) %>%
    mutate(ID = coyote_ids[i], 
           t = NA,
           states = NA,
           daynight = NA,
           sex = NA,
           Used = 0) -> availables[[i]]  
}

# Then combine individual lists into one data.frame
# Then merge with used coyote GPS points data frame
coyote_rest <- availables %>% 
  do.call(rbind,.) %>% 
  rbind(coyote_rest, .) %>% 
  mutate(elev = raster::extract(envt[[1]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         tri = raster::extract(envt[[2]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         veg = raster::extract(envt[[3]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         dev_dist = raster::extract(envt[[4]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         road_dist = raster::extract(envt[[5]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         crop_dist = raster::extract(envt[[6]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE))

# Check that we have a 5:1 ratio of used:available points
table(coyote_rest$ID,coyote_rest$Used)
```

### Foraging top models

### Calculate Forage Home Ranges (Kernel Utilization Distribution)
```{r}
#create object for calculating HRs
coyote_sp <- coyote %>%
   filter(states == 2) %>% #filter to rest only
   dplyr::select(ID,x,y) %>% #select data 
   st_as_sf(., coords = c("x","y"), crs = "+init=epsg:4326") %>%  #make spatial object
   st_transform("+init=epsg:26910") %>% #transform back to UTMs
   as("Spatial") #transform back to sp object for KUDs

#create Kernel Utilization Distributions 
kud_all <- adehabitatHR::kernelUD(coyote_sp, h = 'href')
coyote_hr <- adehabitatHR::getverticeshr(kud_all, percent = 95 , unout = "km2")

#HR area (km2)
kernel.area(kud_all, percent = 95) %>% 
  pivot_longer(1:8,names_to = "ID", values_to = "Area (km^2)")
```


### Sample available points in Home Ranges
```{r}
coyote_forage <- coyote %>% 
  filter(states == 2) %>% #filter to rest only
  mutate(Used = 1) %>% #add column to indicate used points
   st_as_sf(., coords = c("x","y"), crs = "+init=epsg:4326") %>%  #make spatial object
   st_transform("+init=epsg:26910")  #transform back to UTMs

coyote_ids <- unique(coyote_forage$ID) # Extract a list of IDs
availables <- list() #empty list to store available pts by ID

# Randomly sample available points from within the home range of each individual
for(i in 1:length(coyote_ids)){
    st_sample(st_as_sf(coyote_hr)[i,], 5*nrow(filter(coyote_forage, ID == coyote_ids[i]))) %>%  #5:1 ratio
    st_sf(geometry = .) %>%
    mutate(ID = coyote_ids[i], 
           t = NA,
           states = NA,
           daynight = NA,
           sex = NA,
           Used = 0) -> availables[[i]]  
}

# Then combine individual lists into one data.frame
# Then merge with used coyote GPS points data frame
coyote_forage <- availables %>% 
  do.call(rbind,.) %>% 
  rbind(coyote_forage, .) %>% 
  mutate(elev = raster::extract(envt[[1]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         tri = raster::extract(envt[[2]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         veg = raster::extract(envt[[3]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         dev_dist = raster::extract(envt[[4]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         road_dist = raster::extract(envt[[5]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         crop_dist = raster::extract(envt[[6]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE))

# Check that we have a 5:1 ratio of used:available points
table(coyote_forage$ID,coyote_forage$Used)
```

### Traveling top models

### Calculate Travel Home Ranges (Kernel Utilization Distribution)
```{r}
#create object for calculating HRs
coyote_sp <- coyote %>%
   filter(states == 3) %>% #filter to rest only
   dplyr::select(ID,x,y) %>% #select data 
   st_as_sf(., coords = c("x","y"), crs = "+init=epsg:4326") %>%  #make spatial object
   st_transform("+init=epsg:26910") %>% #transform back to UTMs
   as("Spatial") #transform back to sp object for KUDs

#create Kernel Utilization Distributions 
kud_all <- adehabitatHR::kernelUD(coyote_sp, h = 'href')
coyote_hr <- adehabitatHR::getverticeshr(kud_all, percent = 95 , unout = "km2")

#HR area (km2)
kernel.area(kud_all, percent = 95) %>% 
  pivot_longer(1:8,names_to = "ID", values_to = "Area (km^2)")
```


### Sample available points in Home Ranges
```{r}
coyote_travel <- coyote %>% 
  filter(states == 3) %>% #filter to rest only
  mutate(Used = 1) %>% #add column to indicate used points
   st_as_sf(., coords = c("x","y"), crs = "+init=epsg:4326") %>%  #make spatial object
   st_transform("+init=epsg:26910")  #transform back to UTMs

coyote_ids <- unique(coyote_travel$ID) # Extract a list of IDs
availables <- list() #empty list to store available pts by ID

# Randomly sample available points from within the home range of each individual
for(i in 1:length(coyote_ids)){
    st_sample(st_as_sf(coyote_hr)[i,], 5*nrow(filter(coyote_travel, ID == coyote_ids[i]))) %>%  #5:1 ratio
    st_sf(geometry = .) %>%
    mutate(ID = coyote_ids[i], 
           t = NA,
           states = NA,
           daynight = NA,
           sex = NA,
           Used = 0) -> availables[[i]]  
}

# Then combine individual lists into one data.frame
# Then merge with used coyote GPS points data frame
coyote_travel <- availables %>% 
  do.call(rbind,.) %>% 
  rbind(coyote_travel, .) %>% 
  mutate(elev = raster::extract(envt[[1]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         tri = raster::extract(envt[[2]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         veg = raster::extract(envt[[3]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         dev_dist = raster::extract(envt[[4]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         road_dist = raster::extract(envt[[5]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE),
         crop_dist = raster::extract(envt[[6]], as(., "Spatial")) %>% scale(., center = TRUE, scale = TRUE))

# Check that we have a 5:1 ratio of used:available points
table(coyote_travel$ID,coyote_travel$Used)
```


### Behavior Mods
```{r}
#import best full model 
load(here("data_objects","full_mod.Rdata"))

rest_mod <- glmer(Used ~ tri + veg + road_dist + dev_dist + crop_dist + (1|ID), 
                  data = coyote_rest, family=binomial(link="logit")) #drop elev (correlated w/ crop)

forage_mod <- glmer(Used ~ tri + veg + road_dist + dev_dist + crop_dist + (1|ID),
                  data = coyote_forage, family=binomial(link="logit")) #drop elev (correlated w/ crop)

travel_mod <- glmer(Used ~ tri + veg + road_dist + dev_dist + crop_dist + (1|ID),
                  data = coyote_travel, family=binomial(link="logit")) #drop elev (correlated w/ crop)

model_performance(full_mod)
model_performance(rest_mod)
model_performance(forage_mod)
model_performance(travel_mod)
```

```{r}
#Best model
jtools::plot_summs(full_mod, rest_mod, forage_mod, travel_mod, scale = TRUE, 
           model.names = c("Full", "Resting", "Foraging", "Traveling"),
           coefs = c("Ruggedness" = "tri",
                     "Vegetation Type" = "veg",
                     "Distance to Road" = "road_dist",
                     "Distance to Development" = "dev_dist",
                     "Distance to Agriculture" = "crop_dist"),
           colors = c("black","#CC9900","#339999","#006633")) 
```