---
title: "2. Pre-process Environmental Covariates"
author: "Amy Van Scoyoc"
date: "11/21/2022"
output: github_document
---

# Purpose 

This section contains the code to reclassify environmental covariates for use in the resource selection functions (RSFs). The initial `.tif` files were obtained from the sources listed. The layers were cropped to the study area (Sanel Valley, Mendocino County, California, USA) using Google Earth Engine before importing for reclassification here.   

## Load Libraries
```{r warning=F, message=F}
library(here) #for setting working directory reproducibly
library(tidyverse) #data tidying
library(sf) #modifying spatial attributes 
library(raster) #reclassify environmental covariates
library(tmap) #raster plots
```

This function reclassifies the layers, converts to a distance raster (if desired), and writes to file
```{r}
mat_rec <- function(layer, reclass_vec, distance=F, filename){
  
  mat <- matrix(reclass_vec, ncol = 3, byrow = TRUE) #create matrix
  rec <- reclassify(layer, mat) #reclassify values 

  if(distance==T){
    rec <- gridDistance(rec, origin = 1) #make into distance raster if desired
  }
  
  if(missing(filename)) { #don't write to file if missing
    rec #return layer
  } else { #otherwise write to file if specified
        writeRaster(rec, filename = here("data_covariates", filename), 
                    format = "GTiff", overwrite = T) 
        rec #return layer
    }
}
```

### Land Cover
_*Source:*_ 
```{r}
#import landcover layer
lc <- raster(here("data_covariates", "HREC_landcover_extended.tif"))

#reclassify to developed (0), cultivated (1), and natural (2)
lc_vec <- c(0, 20, NA,  # water set to NA 
            20, 21, 1,  # developed open space set cultivated as 1s
            21, 30, 0,  # developed / hardscape set to 0s
            30, 31, NA, # barren set to NA
            31, 80, 2,  # natural habitats set to 2s 
            80, 89, 1,  # cultivated space set to 1s
            89, Inf, NA)# wetlands set to NA 

lc_reclass <- mat_rec(lc, lc_vec, filename = "HREC_landcover_reclass.tif") #reclassify 
```

### Roads
_*Source:*_ 
```{r}
#import roads layer
road <- raster(here("data_covariates", "HREC_roads_extended.tif"))

#reclassify to roads and NAs
road_vec <- c(-1, 50, NA, # non-road set to NA 
              50, Inf, 1) # all roads set to 1 

road_reclass <- mat_rec(road, road_vec, T, filename = "HREC_roads_reclass.tif") #reclassify 
```

### Water
_*Source:*_ 
```{r}
#import water layer
water <- raster(here("data_covariates", "HREC_water_extended.tif"))

#reclassify to water and NAs
water_vec <- c(-1, 0, NA, #non-water set to NA 
                0, Inf, 1) #all water set to 1 

water_reclass <- mat_rec(water, water_vec, T, filename = "HREC_water_reclass.tif") #reclassify 
```

### Human Footprint 
_*Source:*_ (2015) 
```{r}
#import dog point layer
hf <- raster(here("data_covariates", "HREC_humanfootprint_extended.tif"))

#create distance surface (meters)
hf_dist <- gridDistance(hf, origin = 255)

#write to file
writeRaster(hf_dist, filename = here("data_covariates", "HREC_hf_dist.tif"), 
            format = "GTiff", overwrite = T)
```

## Reclassified Layers
```{r, echo=F}
tmap_arrange(
  tm_shape(lc_reclass) + tm_raster(style = "cont", palette="PuBuGn") +
    tm_layout(legend.show = T, legend.outside = T, panel.show=T, panel.labels='Land Cover'),
  tm_shape(water_reclass) + tm_raster(style = "cont", palette ="-Blues") +
    tm_layout(legend.show = T, legend.outside = T, panel.show=T, panel.labels='Water Distance'),
  tm_shape(road_reclass) + tm_raster(style = "cont", palette ="magma") +
    tm_layout(legend.show = T, legend.outside = T, panel.show=T, panel.labels='Road Distance'), 
  tm_shape(hf_dist) + tm_raster(style = "cont") +
    tm_layout(legend.show = T, legend.outside = T, panel.show=T, panel.labels='Development Distance')
)
```

